### 2024-12-31 doc/thesis/pre_info/acknowledge.tex

\chapter*{Acknowledgements}

I would like to express my gratitude to my thesis advisors, Prof. Ryo Maezono and Associate Prof. Kenta Hongo for their advice, support, and the opportunities they have given to me during my Master's year in JAIST.

The research topic on movement prediction of aperiodic time-series data was brought to me by Assistant Prof. Ichiba Tomohiro, so I would also like to thank him for introducing me to this problem and for the guidance he provided.

This journey would never have begun without the enthusiastic support and guidance from Prof. Le Hoai Bac. Thank you for leading me to JAIST and for everything.

I would like to thank everyone in Maezono and Hongo laboratory, fellow students and the laboratory staff, for helping me out whenever I run into trouble with the administrative documents or health.

Lastly, I wish to thank my parents for their never-ending support, motivation, and phone calls, and for helping me to stay connected with my family and friends at home.