### 2025-01-30 doc/thesis/pre_info/bio.tex

\begin{center}
{\LARGE\bf Biography}
\end{center}
\vspace{20pt}

\begin{table}[h]
\setlength{\tabcolsep}{0.7ex}
\begin{tabular}{lll}
\textbf{Name} && Rohit Sanjay Dahule \\[3ex]
\textbf{Date of Birth} && 23 May 1994 \\[3ex]
\textbf{Educational Attainment} && 2020: M.Tech. in Materials Science \& Engineering \\[0.5ex] 
    && Indian Institute of Technology Gandhinagar, \\[0.5ex]
    && Palaj, Gujarat, India \\[2ex]
    && 2016: B.E. in Mechanical Engineering\\[0.5ex] 
    && Rajiv Gandhi College of Engineering, Research and Technology \\[0.5ex]
    && Chandrapur, Maharashtra, India \\[2ex]

\textbf{Scholarship} && 2021-2025: Monbukagakusho (MEXT) Scholarship \\[2ex]
\textbf{Work Experience} && 2020-2021: Junior Research Fellow \\[0.5ex]
    && Indian Institute of Technology Gandhinagar, \\[0.5ex] 
    && Palaj, Gujarat, India \\[2ex]
\end{tabular}
\end{table}
\vspace{-10pt} 
\noindent
\textbf{Publications}
\begin{enumerate}[label={[\arabic*]}]
    \item \underline{Dahule R.}, Raghav A., Hanindriyo A. T., Hongo K., Maezono R.,
        \& Panda, E. (2021). Surface study of Cu$_2$SnS$_3$ using first‐principles
        density functional theory,  Adv. The. Simu., 4(6), 2000315.
    \item \underline{Dahule R.}, Singh C. C., Hongo K., Maezono R., \& Panda E.
        (2022). Anomalies in the bulk and surface electronic properties of SnS:
        effects of native defects,  J. Mate. Chem. C, 10(14), 5514-5525.
    \item Song P., Durajski A, Hou Z., Ghaffar A, \underline{Dahule R.}, Szczesniak R.,Hongo K.,
        \& Maezono R., (2024) (La,Th)H$_{10}$: Potential high-T$_c$ (242 K) 
        superconductors stabilized thermodynamically below 200 GP. 
        J. Phys. Chem. C 2024, 128, 6, 2656–2665.
    \item Bandi H., Kakarla A., \underline{Dahule R.}, Maezono R., Narsimulu D.,
         Shanthappa R. \& Yu J. S.,
         Nickel vanadate cathode induced in-situ phase transition for improved zinc storage
         by low migration barrier and Zn$^{2+}$/H$^+$ co-insertion mechanism.
         small 2024, 2408568.
    \item \underline{Dahule R.}, Gao, B., Hongo, K., Panda E., Ma, Y. \& Maezono R., 
        Stability and Electronic Properties of SnS/ZnS Interfaces: 
        A First-Principles Investigation, 
        (under review for J. Phys. Chem. C)
    \item \underline{Dahule R.}, Oqumhula K., Maezono, R., \& Hongo, K., 
        Physics-informed data-driven discovery of high thermal conductive polymer crystals. 
        (under review for ACS Appl. Polym. Mater.)
    \item Bandi H., \underline{Dahule R.}, Kakarla A., Maezono R., Narsimulu D., 
        Akram S. W.,  Shanthappa R. \& Yu J. S., 
        New insights into the green synthesis of CoMoO$_{4}$ nanorods decorated with 
        silver nanoparticles for high-performance Li$^+$ storage anode 
        (under review for J. Mate. Chem. A)
\end{enumerate}